---
title: "simlulation-exercise"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this project, I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. I will compare the sample mean and variance with the theoretical mean and variance. I will plot the distribution and explain how it is approximately normal.

## Simulations

```{r}
library(ggplot2)

# Set constants provided in the assignment instruction
lambda <- 0.2
number_of_exponetials <- 40
number_of_simulations <- 1000

set.seed(0)

# Run the simulation and calculate the means
exponential_matrix = matrix(data=rexp(number_of_exponetials * number_of_simulations, lambda), nrow=number_of_simulations)
exponential_means = data.frame(means=apply(exponential_matrix, 1, mean))

# plot the means for comparison with the Central Limit Theorem 
ggplot(data = exponential_means, aes(x = means)) + geom_histogram(binwidth=0.2)
```

## Compare the sample mean with the theoretical mean


```{r}
theoretical_mean <- 1/lambda
theoretical_mean
```

```{r}
sample_mean <- mean(exponential_means$means)
sample_mean
```

They are pretty much the same.

## Compare the sample variance with the theoretical variance

```{r}
theoretical_variance <- (1/lambda/sqrt(number_of_exponetials))^2
theoretical_variance
```

```{r}
theoretical_variance <- sd(exponential_means$means)
theoretical_variance
```

The difference here is larger than the one regarding the mean, but they are still pretty close.

## Compare the sample distribution with the one of the Central Limit Theorem

```{r}
ggplot(data = exponential_means, aes(x = means)) + 
  geom_histogram(aes(y=..density..), alpha=0.1, binwidth=0.2) + 
  stat_function(fun = dnorm, args = list(mean = theoretical_mean, sd = sqrt(theoretical_variance)), colour = "red", size=1) + 
  geom_density(colour="blue", size=1)
```

The two graphs overlap quite well with each other.
